---
title: "Data Wrangling"
author: "Monique Parisi"
date: "19/06/2019"
output: html_document
---

Loading the tidyverse
```{r}
library(tidyverse)
```

Load data 
```{r}
#loading the gapminder dataset from github
gapminder <- read_csv("https://raw.githubusercontent.com/OHI-Science/data-science-training/master/data/gapminder.csv")
```

```{r}
head(gapminder)
tail(gapminder)

#str() stands for structure 
str(gapminder)

#column names of gapminder
names(gapminder)
ncol(gapminder)
```

Look at individual dataframe variables 
```{r}
#show me the lifeExp column
gapminder$lifeExp
gapminder$continent
```

Learning the 'filter()' function
```{r}
#filtering data to find life expectancey under 29
filter(gapminder, lifeExp <29)

#just look at mexico
filter(gapminder, country == "Mexico", year == "2002")

```

Average life expectancy in Brazil between 1987 and 2007
```{r}
brazil_life_exp <- filter(gapminder, country == "Brazil", year > 1987)
brazil_life_exp
mean (brazil_life_exp$lifeExp)

```

Learning the 'select()' function
```{r}
#these two lines do the same thing
select(gapminder, year, country, lifeExp)
#drop columns by using a -
select(gapminder, -continent, - gdpPercap)
```

Using 'filter()' and 'select()' together
```{r}
gap_cambodia <- filter(gapminder, country == "Cambodia")
gap_cambodia2 <- select(gap_cambodia, -continent, -lifeExp)
```

Pipe operator links functions together 
cmmd+shift+m ('and then' take whatever is on the left and feed it to the right)
```{r}
gapminder %>% head()
head(gapminder)

#these two pipes do the same things
gap_cambodia <- filter(gapminder, country == "Cambodia") %>%  #and then
  select(-continent, -lifeExp)

gap_cambodia <- gapminder %>% 
  filter(country == "Cambodia") %>% 
  select(-continent, -lifeExp)
```

Learn 'mutate()' which adds a new column 
```{r}
#these two do the same thing
gapminder %>% 
  mutate(gdp = pop * gdpPercap) #before the = is the new column name

mutate(gapminder, gdp = pop * gdpPercap)

```

Calculate the population in thousands for all Asian countries in the year 2007 and add it as a new column
```{r}
gapminder %>%
  filter(continent == "Asia",
         year == 2007) %>% 
  mutate(pop_thousands = pop/1000) %>% 
  select(country, year, pop_thousands)
```

Learning the 'group_by()' function
```{r}
gapminder %>% 
  filter(year == 2002) %>% 
  group_by(continent) %>% 
  mutate(cont_pop = sum(pop))
```

Add in 'summarize()'
```{r}
gapminder %>% 
  filter(year == 2002) %>% 
  group_by(continent) %>% 
  summarize(cont_pop = sum(pop))
```

```{r}
gapminder %>% 
  group_by(continent, year) %>% 
  summarize(cont_pop = sum(pop)) %>% 
  ungroup() #good practice to ungroup after you do a grouping variable
```

'arrange()' arranges columns
```{r}
gapminder %>% 
  group_by(continent, year) %>% 
  summarize(cont_pop = sum(pop)) %>% 
  arrange(year)
```

What is the max GDP per continent across all years?
```{r}
gapminder %>% 
  mutate(gdp = pop * gdpPercap) %>% 
  group_by(continent) %>% 
  mutate(max_gdp = max(gdp)) %>% 
  filter(gdp == max_gdp)
```

