---
title: "Graphics with ggplot2"
author: "Monique Parisi"
date: "19/06/2019"
output: html_document
---

# Learning gglot2

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load the tidyverse

```{r}
library(tidyverse)
```

# Load data

```{r data}
ohi_data <- read_csv("https://raw.githubusercontent.com/OHI-Science/data-science-training/master/data/OHI_global_data.csv")
```
```{r}
head(ohi_data)
```

```{r}
summary(ohi_data)
```

```{r}
table(ohi_data$georegion_one)
# the $ sign pulls a collumn
```

```{r}
dim(ohi_data)
```
# Create ggplot 
```{r}
ggplot(data = ohi_data, aes(x = georegion_one, y = OHI_score)) + 
  geom_jitter(width=0.2)
```
ggplot only works with data frames
aes : determines x and y variables 
+ : used to add layers 
geom : added points 
geom tells R what kind of plot 
jitter adds and random variation in the x axis 



```{r Boxplot}
ggplot(data = ohi_data, aes (x = georegion_one)) +
  geom_bar()
```

```{r Histogram}
ggplot(data = ohi_data, aes(x = HDI)) + 
  geom_histogram()
```

```{r Multiple geoms}
ohi_summary <- read_csv ("https://raw.githubusercontent.com/OHI-Science/data-science-training/master/data/OHI_scores_georegion_summary.csv")

ggplot(data=ohi_summary, aes(x = georegions, y = OHI_score_average)) +
  geom_bar(stat="identity") +
  geom_jitter(data = ohi_data, aes(x = georegion_one, y = OHI_score))
```

```{r Adding colours}
ggplot(data = ohi_data, aes(y=OHI_score, x = HDI, color=georegion_one)) +
  geom_point()
ggplot(data = ohi_data) +
  geom_point(aes(y = OHI_score, x = HDI, color=georegion_one))  
```
You can add colous in the ggplot or the geom but when you're doing more layers you should add it in the geom
Anything in aes has to link to a column in the data 

```{r Adding third variable}
ggplot(data=ohi_data, aes(x = OHI_score, y = HDI, size = coastal_pop, colour = coastal_pop)) +
  geom_point()
```

```{r Adding third discrete variable}
ggplot (data=ohi_data, aes(x=OHI_score, y = HDI, colour = georegion_one)) +
  geom_point()

ggplot (data=ohi_data, aes(x=OHI_score, y = HDI, shape = georegion_one)) +
  geom_point()
```

```{r Themes}
ggplot(data=ohi_data, aes(x = OHI_score, y = HDI)) +
  geom_point() +
  theme_bw()
```
theme_bw() : removes grey background

```{r Creating your own themes}
source('https://raw.githubusercontent.com/OHI-Science/ohiprep/master/src/R/scatterTheme.txt')

ggplot(data = ohi_data, aes(x = OHI_score, y = HDI)) + 
  geom_point() + 
  scatterTheme
```

```{r Change labels}
ggplot(data = ohi_data, aes(x = OHI_score, y = HDI, colour = georegion_one)) + 
  geom_point() + 
  labs(y = "OHI score, 2017", 
       x = "Human Development Index",
       title = "Human development is good for oceans!",
       colour = "Georegions")
```

```{r Change colours}
install.packages("RColorBrewer")
library(RColorBrewer)

display.brewer.all()

my_palette <- brewer.pal(n = 9, "YlOrRd")  #will choose 9 colours from the palette 

ggplot(data=ohi_data, aes(x = OHI_score,
                         y = OHI_trend,
                         color = HDI)) +
  geom_point(size = 3) +
  scale_color_gradientn(colors = c("blue", "red"))

ggplot(data=ohi_data, aes(x = OHI_score,
                         y = OHI_trend,
                         color = georegion_one)) +
  geom_point(size = 3) +
  scale_color_manual(values = my_palette)

```
